---
title: "Mini Project 1"
author: "Jin Kuang"
---

## Task 1
```{r}
if(!dir.exists(file.path("data", "mp01"))){
    dir.create(file.path("data", "mp01"), showWarnings=FALSE, recursive=TRUE)
}

GLOBAL_TOP_10_FILENAME <- file.path("data", "mp01", "global_top10_alltime.csv")

if(!file.exists(GLOBAL_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-global.tsv", 
                  destfile=GLOBAL_TOP_10_FILENAME)
}

COUNTRY_TOP_10_FILENAME <- file.path("data", "mp01", "country_top10_alltime.csv")

if(!file.exists(COUNTRY_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-countries.tsv", 
                  destfile=COUNTRY_TOP_10_FILENAME)
}
```

```{r}
if(!require("tidyverse")) install.packages("tidyverse")
library(readr)
library(dplyr)

GLOBAL_TOP_10 <- read_tsv(GLOBAL_TOP_10_FILENAME)
```

## Task 2
```{r}
GLOBAL_TOP_10 <- GLOBAL_TOP_10 |>
    mutate(season_title = if_else(season_title == "N/A", NA_character_, season_title))
```

## Task 3
```{r}
COUNTRY_TOP_10 <- read_tsv(COUNTRY_TOP_10_FILENAME,
                          na = "N/A")
```


## Task 4
### Q1
```{r}
library(dplyr)

country_n <- COUNTRY_TOP_10 |>
  distinct(country_name) |>
  count()

country_n
```

Netflix operates in `{r} country_n` different countries.


### Q2
```{r}
library(DT)
library(stringr)

GLOBAL_TOP_10 |>
  filter(category == "Films (Non-English)") |>    
  group_by(show_title) |> 
  summarise(`Cumulative Weeks` = cumulative_weeks_in_top_10) |> 
  arrange(desc(`Cumulative Weeks`)) |> 
  rename(`Show Title` = show_title) |>
  head(20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) 


```

Therefore, "All Quiet on the Western Front" has spent the most cumulative weeks 
in the global top 10, and for 23 weeks.

### Q3
```{r}
GLOBAL_TOP_10 |>
  filter(category %in% c("Films (English)", "Films (Non-English)")) |>                 
  group_by(show_title) |>
  summarise(`Runtime (Minutes)` = round(60 * runtime)) |>
  arrange(desc(`Runtime (Minutes)`)) |>
  rename(`Show Title` = show_title) |>
  head(20) |>
  datatable(options = list(searching = FALSE, info = FALSE))
```

Therefore, the longest film (English or non-English) to have ever appeared in the Netflix global Top 10
is "Pushpa 2: The Rule (Reloaded Version)", which lasts for 224 minutes.

### Q4
```{r}
GLOBAL_TOP_10 |>
  group_by(category, show_title) |>
  summarise(`Total Hours of Global Viewership` = sum(weekly_hours_viewed)) |>
  arrange(desc(`Total Hours of Global Viewership`)) |>
  filter(`Total Hours of Global Viewership` == first(`Total Hours of Global Viewership`)) |>
  rename(`Show Title` = show_title,
         Category = category) |>
  datatable(options = list(searching = FALSE, info = FALSE))
```

Thus, Squid Game has the most total hours of global viewership for non-English TV; 
Stranger Things has the most total hours of global viewership for English TV; 
Red Notice has the most total hours of global viewership for English films; 
Society of the Snow has the most total hours of global viewership for non-English films.

### Q5
```{r}
COUNTRY_TOP_10 |>
  filter(category == "TV") |>    
  group_by(show_title, country_name) |> 
  summarise(`Run Time` = cumulative_weeks_in_top_10) |> 
  arrange(desc(`Run Time`)) |> 
  rename(`Show Title` = show_title,
         `Country Name` = country_name) |>
  head(20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) 
```

Thus, Money Heist had the longest run in a country’s Top 10. It occurred in Pakistan and it lasted
for 127 weeks.

### Q6
```{r}
COUNTRY_TOP_10 |>
  distinct(country_name, week) |>            
  count(country_name) |>
  rename(`Number of Weeks` = n) |>
  filter(`Number of Weeks` < 200) |>             
  datatable(options = list(searching = FALSE, info = FALSE))
```

Thus, this country that Netflix provides less than 200 weeks of service history is Russia.

```{r}
COUNTRY_TOP_10 |>
  filter(country_name == "Russia") |>
  group_by(country_name) |> 
  summarise(`Week Date` = week) |>
  arrange(desc(`Week Date`)) |> 
  rename(`Country Name` = country_name) |>
  head(20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) 
```

Thus, Netflix cease operations in that country on 2022-02-27.

### Q7
```{r}
total_viewership <- GLOBAL_TOP_10 |>
  filter(show_title %in% c("Squid Game", "Squid Game: The Challenge")) |>
  summarise(`Total Viewership` = sum(weekly_hours_viewed))

total_viewership
```

Thus, the total viewership of the TV show Squid Game is `{r} total_viewership` hours.

### Q8
```{r}
library(lubridate)

GLOBAL_TOP_10 |>
  filter(show_title == "Red Notice", year(week) == 2021) |>
  select(week, show_title, weekly_views) |>
  rename(`Week` = week,
         `Show Title` = show_title,
         `Weekly Views` = weekly_views) |>
  datatable(options = list(searching = FALSE, info = FALSE)) 
```

Since from the above table, we find that the weekly views column are null, so we 
instead use the weekly hours viewed as our final answer.

```{r}
views_n <- GLOBAL_TOP_10 |>
  filter(show_title == "Red Notice", year(week) == 2021) |>
  summarise(`Number of Views` = round(sum(weekly_hours_viewed)) / ((60 + 58) / 60))

views_n
```

Thus, there are approximately `{r} views_n` views received in 2021

### Q9
```{r}

```



```{r}
COUNTRY_TOP_10 |>
  filter(country_name == "United States", category == "Films") |>
  group_by(show_title) |>
  arrange(week) |>
  summarise(
    `First Rank`  = first(weekly_rank),
    `Eventually Hit Number 1?` = any(weekly_rank == 1),
    `Week That Hit Number 1` = if (any(weekly_rank == 1))
      week[weekly_rank == 1]
    else NA) |>
  filter(`Eventually Hit Number 1?` == TRUE, `First Rank` != 1) |>
  arrange(desc(`Week That Hit Number 1`)) |>
  rename(`Show Title` = show_title) |>
  datatable(options = list(searching = FALSE, info = FALSE)) 
```

Films that first appeared on the Top 10 chart but then became more popular and 
eventually hit Number 1 are listed above. The most recent film to pull this off is
"KPop Demon Hunters".

```{r}
films_n <- COUNTRY_TOP_10 |>
  filter(country_name == "United States", category == "Films") |>
  group_by(show_title) |>
  arrange(week) |>
  summarise(
    `First Rank`  = first(weekly_rank),
    `Eventually Hit Number 1?` = any(weekly_rank == 1),
    `Week That Hit Number 1` = if (any(weekly_rank == 1))
      week[weekly_rank == 1]
    else NA) |>
  filter(`Eventually Hit Number 1?` == TRUE, `First Rank` != 1) |>
  arrange(desc(`Week That Hit Number 1`)) |>
  distinct(show_title) |>
  nrow()

films_n
```

Thus, there are a total of `{r} films_n` Films reached Number 1 in the US but did not originally debut there

### Q10
```{r}
COUNTRY_TOP_10 |>
  filter(category == "TV") |>
  group_by(show_title, season_title) |>
  arrange(week) |>
  filter(week == first(week)) |>
  summarise(`Number of Country` = length(unique(country_name))) |>
  arrange(desc(`Number of Country`))
```

Thus, Emily in Paris: Season 2 hits the top 10 in the most countries in its debut week.
It charted for 94 countries.  










```{r}
library(stringr)
format_titles <- function(df){
    colnames(df) <- str_replace_all(colnames(df), "_", " ") |> str_to_title()
    df
}

GLOBAL_TOP_10 |> 
    format_titles() |>
    head(n=20) |>
    datatable(options=list(searching=FALSE, info=FALSE)) |>
    formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```

```{r}
GLOBAL_TOP_10 |> 
    select(-season_title) |>
    format_titles() |>
    head(n=20) |>
    datatable(options=list(searching=FALSE, info=FALSE)) |>
    formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```

```{r}
GLOBAL_TOP_10 |> 
    mutate(`runtime_(minutes)` = round(60 * runtime)) |>
    select(-season_title, 
           -runtime) |>
    format_titles() |>
    head(n=20) |>
    datatable(options=list(searching=FALSE, info=FALSE)) |>
    formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```

